---
title: "week2-rexcercise"
output: html_document
---

```{r}
library(readr)        # to import tabular data (e.g. csv)
library(dplyr)        # to manipulate (tabular) data
library(ggplot2)      # to visualize data
library(sf)           # to handle spatial vector data
library(terra)        # To handle raster data
library(lubridate)    # To handle dates and times


wildschwein_BE <- read_delim("wildschwein_BE_2056.csv",",") # adjust path

wildschwein_BE <- st_as_sf(wildschwein_BE, coords = c("E", "N"), crs = 2056, remove = FALSE)
```

#####Task 2: Overview

### Calculate time difference
```{r}
wildschwein_BE<-wildschwein_BE %>%
  group_by(TierID) %>%
  mutate(timelag=as.integer(difftime(lead(DatetimeUTC), DatetimeUTC,units="secs")))
```
### How many individuals were tracked? 
```{r}
ggplot(wildschwein_BE, mapping=aes(DatetimeUTC, TierID))+
  geom_line()+
  labs(
    x="TierID",
    y="DatetimeUTC"
    )
````
Solution: Three individuals were tracked.

####Were they tracked conurrently or sequentially?
```{r}
ggplot(data=wildschwein_BE, mapping=aes(DatetimeUTC, timelag,colour=TierID))+
  geom_line()+
  labs(
    x="timelag",
    y="DatetimeUTC"
  )
````

Solution:They were tracked concurrently.

####For how long were the individuals tracked? Are there gaps?
```{r}
wildschwein_BE %>%
  group_by(TierID)%>%
  summarise(
  difftime(max(DatetimeUTC), min(DatetimeUTC), units = "days"))
```
#Solution: Individual 002A was tracked for 338 days, 016A for 234 days and 018 for 261 days.


####What is the temporal sampling interval between the locations?
```{r}
wildschwein_BE%>%
  group_by(TierID)%>%
  summarise(median(timelag,na.rm=TRUE))

ggplot(data=wildschwein_BE,mapping=aes(x=timelag))+
  geom_histogram(binwidth =10)+
  scale_y_log10()+
  xlim(0,15000)+
  theme_classic()
```
#Solution: On average every individual was sampled every 900 seconds. There are some gaps where it wasn't sampled for more than 10'000 seconds.
  

#####Task 3: Movement parameters - Speed 
  
####Calculate steplength
```{r}
wildschwein_BE <- wildschwein_BE %>%
  group_by(TierID) %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
````
####Calculate speed
```{r}
wildschwein_BE <- wildschwein_BE %>%
  mutate(speed=steplength/timelag)
````
#Solution: I get speed unit m/s

#####: Task 4: Cross-scale movement analysis

####Import new dataset
```{r}
caro <- read_delim("caro60.csv",",")
caro <- st_as_sf(caro, coords = c("E", "N"), crs = 2056, remove = FALSE)
````
####Select every 3rd, 6th and 9th row
```{r}
caro_3 <- caro[seq(1,nrow(caro),by=3),]
caro_6 <- caro[seq(1,nrow(caro),by=6),]
caro_9 <- caro[seq(1,nrow(caro),by=9),]
  

nrow(caro)
nrow(caro_3)
nrow(caro_6)
nrow(caro_9)  
````
  
####Calculate timelag for each dataset
```{r}
caro<-caro%>%
  mutate(timelag=as.integer(difftime(lead(DatetimeUTC), DatetimeUTC,units="secs")))
  
caro_3<-caro_3%>%
  mutate(timelag=as.integer(difftime(lead(DatetimeUTC), DatetimeUTC,units="secs")))
  
caro_6<-caro_6%>%
  mutate(timelag=as.integer(difftime(lead(DatetimeUTC), DatetimeUTC,units="secs")))
  
caro_9<-caro_9%>%
  mutate(timelag=as.integer(difftime(lead(DatetimeUTC), DatetimeUTC,units="secs")))
````
####Calculate steplength for each dataset
```{r}
caro <- caro %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
  
caro_3 <- caro_3 %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
  
caro_6<- caro_6 %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
  
caro_9 <- caro_9 %>%
  mutate(steplength = sqrt((E- lead(E,1))^2 + (N -lead(N,1))^2))
````
####Calculate speed for each dataset
```{r}
caro <- caro %>%
  mutate(speed=steplength/timelag)
  
caro_3 <- caro_3 %>%
  mutate(speed=steplength/timelag)
  
caro_6 <- caro_6 %>%
  mutate(speed=steplength/timelag)
  
caro_9 <- caro_9 %>%
  mutate(speed=steplength/timelag)
````

####Visual comparison of original and 3 minutes resampled data
```{r}
ggplot(data=caro,mapping=aes(x=E, y=N))+
  geom_path(data=caro,mapping=aes(color="1 minute"))+
  geom_path(data = caro_3, mapping=aes(color="3 minutes"))+
  geom_point(data=caro,mapping=aes(color="1 minute"))+
  geom_point(data = caro_3, mapping=aes(color="3 minutes"))+
  theme_light()+
  labs(color="Trajectory", title = "Comparison original vs 3 minutes resampled")+
  theme(title=element_text(size=8))
````

####Visual comparison of original and 6 minutes resampled data
```{r}
ggplot(data = caro,mapping=aes(x=E, y=N))+
  geom_path(data= caro,mapping=aes(color="1 minute"))+
  geom_path(data = caro_6, mapping=aes(color="6 minutes"))+
  geom_point(data=caro,mapping=aes(color="1 minute"))+
  geom_point(data = caro_6, mapping=aes(color="6 minutes"))+
  theme_light()+
  labs(color="Trajectory", title = "Comparison original vs 6 minutes resampled")+
  theme(title=element_text(size=8))
````
  

####Visual comparison of original and 9 minutes resampled data
```{r}
ggplot(data = caro,mapping=aes(x=E, y=N))+
  geom_path(data= caro,mapping=aes(color="1 minute"))+
  geom_path(data = caro_9, mapping=aes(color="9 minutes"))+
  geom_point(data=caro,mapping=aes(color="1 minute"))+
  geom_point(data = caro_9, mapping=aes(color="9 minutes"))+
  theme_light()+
  labs(color="Trajectory", title = "Comparison original vs 9 minutes resampled")+
  theme(title=element_text(size=8))
````

####Visual comparison of speeds at different sampling intervals
```{r}
ggplot(data = caro,mapping=aes(x=DatetimeUTC, y=speed))+
  geom_line(data=caro, mapping=aes(colour="1 minute"))+
  geom_line(data=caro_3, mapping=aes(colour="3 minute"))+
  geom_line(data=caro_6, mapping=aes(colour="6 minute"))+
  geom_line(data=caro_9, mapping=aes(colour="9 minute"))+
  theme_light()+
  labs(color="Trajectory", title = "Comparing derived speed at different sampling intervals")+
  theme(title=element_text(size=8))+
  xlab("Time")+
  ylab("Speed (m/s)")
````
#####Task5: Deriving movement parameters - rolling window functions
```{r}
install.packages("zoo")
library(zoo)

example <- rnorm(10)
example
rollmean(example,k = 3,fill = NA,align = "left")

rollmean(example,k = 4,fill = NA,align = "left")
````

####Testing different rolling windows
```{r}
caro_win <- caro
caro_win$k2 <- rollmean(caro_win$speed, k=2, fill = NA, align = "left")
caro_win$k4 <- rollmean(caro_win$speed, k=4, fill = NA, align = "left")
caro_win$k6 <- rollmean(caro_win$speed, k=6, fill = NA, align = "left")
caro_win$k8 <- rollmean(caro_win$speed, k=8, fill = NA, align = "left")
caro_win$k10 <- rollmean(caro_win$speed, k=10, fill = NA, align = "left")
````

####Visualisation different rolling windows
```{r}
ggplot(data = caro_win,mapping=aes(x=DatetimeUTC, y=speed))+
  geom_line(alpha=0.5)+
  geom_line(data=caro_win, mapping=aes(y=k2, color="k2"))+
  geom_line(data=caro_win, mapping=aes(y=k4, color="k4"))+
  geom_line(data=caro_win, mapping=aes(y=k6, color="k6"))+
  geom_line(data=caro_win, mapping=aes(y=k8, color="k8"))+
  geom_line(data=caro_win, mapping=aes(y=k10, color="k10"))+
  theme_light()+
  labs(color="Window size", title = "Comparing derived speed at different window sizes")+
  theme(title=element_text(size=8))+
  xlab("Time")+
  ylab("Speed (m/s)")
````





















